---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# genetic.algo.optimizeR

<!-- badges: start -->

<!-- badges: end -->

The goal of `genetic.algo.optimizeR` is to optimize the function $f(x) = x^2 - 4x + 4$ using a genetic algorithm. The function represents a simple quadratic equation, and the goal is to find the value of $x$ that minimizes the function.

Here's a breakdown of the aim and the results:

**Aim:**

-   Optimize the function $f(x) = x^2 - 4x + 4$ to find the value of $x$ that minimizes the function.

**Results:**

-   **Initial Population**:

    -   We start with a population of three individuals: $x_1 = 1$, $x_2 = 3$, and $x_3 = 0$. <br>

-   **Evaluation**:

    -   We evaluate the fitness of each individual by calculating $f(x)$ for each $x$ value:
        -   $f(1) = 1^2 - 4*1 + 4 = 1$
        -   $f(3) = 3^2 - 4*3 + 4 = 1$
        -   $f(0) = 0^2 - 4*0 + 4 = 4$

-   **Selection**:

    -   We select individuals $x_1$ and $x_2$ as parents for crossover because they have higher fitness.

-   **Crossover and Mutation**:

    -   We perform crossover and mutation on the selected parents to generate offspring: $x_1' = 1$, $x_2' = 3$.

-   **Replacement**:

    -   We replace individual $x_3$ with offspring $x_1'$, maintaining the population size.

After multiple generations of repeating these steps, the genetic algorithm aims to converge towards an optimal or near-optimal solution. In this example, since it's simple and the solution space is small, we could expect the algorithm to converge relatively quickly towards the optimal solution $x = 2$, where $f(x) = 0$.

***Explaining Graph***

``` mermaid
graph TD;
    A[Initialize Population] --> B(Evaluate Fitness);
    B --> C[Selection];
    C --> D{Termination Condition?};
    D -->|Yes| E[End];
    D -->|No| F[Crossover and Mutation];
    F --> G[Replace Population];
    G --> B;
    B -->|Example:| X1(Individual 1: x = 1);
    B --> X2(Individual 2: x = 3);
    B --> X3(Individual 3: x = 0);
    C -->|Example:| Y1(Selected Parents: x = 1, x = 3);
    F -->|Example:| Z1(Offspring 1: x = 1);
    F --> Z2(Offspring 2: x = 3);
```

## Usage

```{r usage}
library(genetic.algo.optimizeR)

# Initialize population
population <- initialize_population(population_size = 3, min = 0, max = 3)
print("Initial Population:")
print(population)

while (TRUE) {
  # Evaluate fitness
  fitness <- evaluate_fitness(population)
  print("Evaluation:")
  print(fitness)
  
  
  # Check if the fitness of every individual is close to zero
  if (all(abs(fitness) <= 0.01)) {
    print("Termination Condition Reached: All individuals have fitness close to zero.")
    break
  }
  
  # Selection
  selected_parents <- selection(population, fitness, num_parents = 2)
  print("Selection:")
  print(selected_parents)
  
  # Crossover and Mutation
  offspring <- crossover(selected_parents, offspring_size = 2)
  mutated_offspring <- mutation(offspring, mutation_rate = 0) #(no mutation in this example)
  print("Crossover and Mutation:")
  print(mutated_offspring)
  
  # Replacement
  population <- replacement(population, mutated_offspring, num_to_replace = 1)
  print("Replacement:")
  print(population)
}

```

The above example illustrates the process of a genetic algorithm, where individuals are selected, crossed over, and replaced iteratively to improve the population towards finding the optimal solution(i.e. fitting population).

***In theory***

1.  Initialize Population:
    -   Start with a population of individuals: X1(x=1), X2(x=3), X3(x=0).\
    (Note: the values are random and the population should be highly diversified)
    -   The space of x value is kept integer type and on range from 0 to 3,for simplification.
```{r}
population <- c(1, 3, 0)
population
```
2.  Evaluate Fitness:
    -   Calculate fitness(`f(x)`) for each individual:

        -   X1: f(1) = 1\^2 - 4\*1 + 4 = 1
        -   X2: f(3) = 3\^2 - 4\*3 + 4 = 1
        -   X3: f(0) = 0\^2 - 4\*0 + 4 = 4

        Coding the function f(x) in R A quadratic function is a function of the form: ax2+bx+c where a≠0

  So for:
  
  $$
  f(x) = x^2 - 4x + 4
  $$
\
In R, we write:

```{r}
a = 1
b = -4
c = 4
f = function(x) {
  a*x^2 + b*x + c
}
```

Plotting the quadratic function f(x) First, we have to choose a domain over which we want to plot f(x).

Let’s try 0 ≤ x ≤ 3:

```{r initial_poputation, out.width="70%", fig.cap=""}
# domain over which we want to plot f(x)
x = seq(from=0, to=4, length.out=100)
# plot f(x)
plot(x, f(x), type = 'l') # type = 'l' plots a line instead of points
# plot the x and y axes
abline(v = 0,h = 0,  col = "skyblue", lty = 3)
points(c(1,3), c(f(1),f(3)), col = "coral1",pch = 8, cex = 1.5, lty = 3)
text(c(1, 3), c(f(1), f(3)), labels = c("(x=1, f(x=1))", "(x=3, f(x=3))"), pos = 3)

points(c(0), c(f(0)), col = "blue",pch = 8, cex = 1.5, lty = 3)
text(c(0), c(f(0)), labels = "(x=0, f(x=0))", pos = 4, font = 2)

```

3.  Selection:
    -   Select parents for crossover:

        -   Y1(x=1), Y2(x=3)

```{r selection, out.width="70%", fig.cap=""}
# plot f(x)
plot(x, f(x), type = 'l') # type = 'l' plots a line instead of points
# plot the x and y axes
points(c(1,3), c(f(1),f(3)), col = "coral1",pch = 8, cex = 2, lty = 3)
  text(c(1, 3), c(f(1), f(3)), labels = c("(x=1, f(x=1))", "(x=3, f(x=3))"), pos = 3, font = 2)
```

4.  Crossover and Mutation:
    -   Generate offspring through crossover and mutation:
        -   Z1(x=1), Z2(x=3) (no mutation in this example)
5.  Replacement:
    -   Replace individuals in the population:
        -   Replace X3 with Z1, maintaining the population size.
6.  Repeat Steps 2-5 for multiple generations until a termination condition is met.

The optimal/fitting individuals *F* of a quadratic equation, in this case the lowest point on the graph of f(x), is: $F(\frac{−b}{2a},f(\frac{−b}{2a}))$

```{r fitting, out.width="70%", fig.cap=""}
find.fitting = function(a, b, c) {
  x_fitting = -b/(2 * a)
  y_fitting = f(x_fitting)
  c(x_fitting, y_fitting)
}
F = find.fitting(a, b, c)
```

Adding the Fitting to the plot:

```{r fitting_plot, out.width="70%", fig.cap=""}
# plot f(x)
plot(x, f(x), type = 'l') # type = 'l' plots a line instead of points
# plot the x and y axes
abline(h = 0)
abline(v = 0)
# add the vertex to the plot
points(x = F[1], y = F[2],
       pch = 18, cex = 2, col = 'red') # pch controls the form of the point and cex controls its size
# add a label next to the point
text(x = F[1], y = F[2],
     labels = "Fitting", pos = 3, col = 'red', font = 10) # pos = 3 places the text above the point
```

***Existing alternative solution***



Finding the x-intercepts of ***f(x)***

The x-intercepts are the solutions of the quadratic equation f(x) = 0; they can be found by using the quadratic formula:

$$
x = \frac{-b \pm \sqrt{b^2 - 4ac}}{2a}
$$

The quantity $b2–4ac$ is called the discriminant:

-   if the discriminant is positive, then *f(x)* has 2 solutions (i.e. x-intercepts).
-   if the discriminant is zero, then *f(x)* has 1 solution (i.e. 1 x-intercept).
-   if the discriminant is negative, then *f(x)* has no real solutions (i.e. does not intersect the x-axis).


```{r function_intercepts}
# find the x-intercepts of f(x)
find.roots = function(a, b, c) {
  discriminant = b^2 - 4 * a * c
  if (discriminant > 0) {
    c((-b - sqrt(discriminant))/(2 * a), (-b + sqrt(discriminant))/(2 * a))
  }
  else if (discriminant == 0) {
    -b / (2 * a)
  }
  else {
    NaN
  }
}
solutions = find.roots(a, b, c)
```

Adding the x-intercepts to the plot:
```{r plot_intercepts, out.width="70%", fig.cap=""}
# plot f(x)
plot(x, f(x), type = 'l') # type = 'l' plots a line instead of points
# plot the x and y axes
abline(h = 0)
abline(v = 0)
# add the x-intercepts to the plot
points(x = solutions, y = rep(0, length(solutions)), # x and y coordinates of the x-intercepts
       pch = 18, cex = 2, col = 'red')
text(x = solutions, y = rep(0, length(solutions)),
     labels = rep("Fitting(x-intercept)", length(solutions)),
      pos = 3, col = 'red', font = 10)
```

