---
title: "Genetic Algorthim Optimization"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(genetic.algo.optimizeR)
library(dplyr)
library(ggplot2)
library(magrittr)
library(ggrepel)
knitr::opts_chunk$set(echo = FALSE, collapse = TRUE)
```


## Topic 1

### Exercise 1

*Here's a simple exercise with code chunks provided for entering or editing the answer.*

Use the the function `initialize_population` to generate number of individuals:

```{r two-plus-two, exercise=TRUE}
population <- genetic.algo.optimizeR::initialize_population(population_size = 10, min = 0, max = 3)
population
```

The generated population contains integers within specified range.

### Exercise 2

Given a poputation of these individuals (1, 3, 0), use the the function `evaluate_fitness` to return correct fitness values of all individuals of the popultation:

```{r evaluate_fitness, exercise=TRUE,exercise.lines = 5}
population = c(1, 3, 0)
fitness <- genetic.algo.optimizeR::evaluate_fitness(population)
fitness
```

The expected fitness of the population should (1, 1, 4) respectively.

```{r view_population_fitness, exercise=TRUE,exercise.lines = 7}
population = c(1, 3, 0) # Done above
fitness <- genetic.algo.optimizeR::evaluate_fitness(population) # Done above
population |>
  cbind(fitness) |>
  data.frame() |>
  magrittr::set_rownames(c("individual_1", "individual_2", "individual_3"))
```


### Exercise 3

*In this next exercise examine the fitness of the given population.*


```{r add-function, exercise=TRUE, exercise.lines = 14}
population = c(1, 3, 0) # Done above
fitness <- genetic.algo.optimizeR::evaluate_fitness(population) # Done above
population |>
  cbind(fitness) |>
  data.frame() |>
  ggplot2::ggplot(aes(x = fitness)) +
  geom_density(fill = "skyblue", color = "steelblue", alpha = 0.7) +
  geom_point(aes(x = fitness, y = 0), color = "red", size = 3) + # Add red points
  geom_vline(xintercept = fitness, linetype = "dashed", color = "red") + # Add vertical line
  ggrepel::geom_text_repel(aes(label = paste0("x=",population)), x = fitness, y = 6, vjust = -0.5, color = "black", size = 5) + # Add text label
  labs(x = "Fitness", y = "Density", 
       title = "Density Plot of Fitness") +
  xlim(0.5, 4.5)
```


## Topic 2

### Exercise 1

*Here's an exercise where the objective is to find the optimal(minimal of `evaluate_fitness`) fitness.*

Recall the initial generated population and fitness(in ascending order):

```{r population_fitness_ascending_order, exercise=TRUE}
population |>
  cbind(fitness) |>
  data.frame() |>
  dplyr::group_by_all() |>
  dplyr::arrange(fitness) # in ascending order based on fitness
```

Select correct number of individuals wit the best fitness: 

```{r selected_parents, exercise=TRUE}
selected_parents <- genetic.algo.optimizeR::selection(population, fitness, num_parents = 2)
selected_parents
```


### Exercise 2

Use crossover to generate correct and defined number of offspring:
```{r offspring, exercise=TRUE}
offspring <- genetic.algo.optimizeR::crossover(selected_parents, offspring_size = 2)
offspring
```

Mutate correctly offspring give a certain rate of mutation:

```{r mutated_offspring, exercise=TRUE}
mutated_offspring <- genetic.algo.optimizeR::mutation(offspring, mutation_rate = 1) # All offspring will be mutated
mutated_offspring
```

Replace correct number of individuals from the previous generation:
```{r replaced_population, exercise=TRUE}
replaced_population <- genetic.algo.optimizeR::replacement(population, offspring, num_to_replace = 1)
replaced_population
```


## Topic 3

### Exercise 1

Optimize this script:

The scirpt below illustrates the process of a genetic algorithm, where individuals are selected, crossed over, and replaced iteratively to improve the population towards finding the optimal solution(i.e. fitting population).

```{r Optimize_this_script, exercise=TRUE, exercise.lines = 40}
print("Initial Population:")
print(population)

generation <- 0  # Initialize generation/reputation counter

while (TRUE) {
  generation <- generation + 1  # Increment generation/reputation count
  
  # Evaluate fitness
  fitness <- evaluate_fitness(population)
  print("Evaluation:")
  print(fitness)
  
  
  # Check if the fitness of every individual is close to zero
  if (all(abs(fitness) <= 0.01)) {
    print("Termination Condition Reached: All individuals have fitness close to zero.")
    break
  }
  
  # Selection
  selected_parents <- selection(population, fitness, num_parents = 2)
  print("Selection:")
  print(selected_parents)
  
  # Crossover and Mutation
  offspring <- crossover(selected_parents, offspring_size = 2)
  mutated_offspring <- mutation(offspring, mutation_rate = 0) #(no mutation in this example)
  print("Crossover and Mutation:")
  print(mutated_offspring)
  
  # Replacement
  population <- replacement(population, mutated_offspring, num_to_replace = 1)
  print("Replacement:")
  print(population)
}

print(paste("Total generations/reputations:", generation))
```

## Topic 4

### Exercise 1

```{r call_libraries4, exercise=TRUE}
library(genetic.algo.optimizeR)
library(dplyr)
library(ggplot2)
```

After multiple generations of repeating the steps(in the Topic 1 and 2), the genetic algorithm will converge towards an optimal or near-optimal solution. In that example, since it's simple and the solution space is small, we could expect the algorithm to converge relatively quickly towards the optimal solution $x = 2$, where $f(x) = 0$.

Now in case case of large population, find the optimal(minimal of `evaluate_fitness`) fitness using the optimize function(from stats Rpackage).

```{r print-limit_optimal_fitting_individual, exercise=TRUE, exercise.eval=TRUE}
population =c(1, 3, 0)
optimal_result <- optimize(genetic.algo.optimizeR::evaluate_fitness, population, maximum = FALSE)
optimal_fitting_individual <- optimal_result$minimum
```

Generate some sample fitness data
```{r print-limit_fitness_data, exercise=TRUE, exercise.eval=TRUE}
population =c(1, 3, 0)
optimal_result <- optimize(genetic.algo.optimizeR::evaluate_fitness, population, maximum = FALSE)
optimal_fitting_individual <- optimal_result$minimum
fitness_data <- rnorm(1000, mean = optimal_fitting_individual, sd = 1)
```


Create a data frame for the density plot

```{r print-limit_df_density, exercise=TRUE}
population =c(1, 3, 0)
optimal_result <- optimize(genetic.algo.optimizeR::evaluate_fitness, population, maximum = FALSE)
optimal_fitting_individual <- optimal_result$objective

fitness_data <- rnorm(1000, mean = optimal_fitting_individual, sd = 1)

df_density <- data.frame(fitness = fitness_data)

df_density |>
  data.frame() |>
  ggplot2::ggplot(aes(x = fitness)) +
  geom_density(fill = "skyblue", color = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = optimal_fitting_individual, linetype = "dashed", color = "red") +
  geom_text(aes(label = paste0("x=",optimal_result$minimum)), x = optimal_result$objective, y = .35, vjust = -0.5, color = "black", size = 5) + # Add text label
  labs(x = "Fitness", y = "Density", 
       title = "Density Plot of Fitness with Optimal Population",
       caption = paste("Optimal Fitting Individual:", round(optimal_result$minimum, 2)))
```

### Quiz

*Here's a section including number of single or multiple choice questions as a quiz.*

Please answer this questions:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

